An Algorithm for Searching Evolving Graphs in Linear Algebraic Language

### 2016-01-22 paper.tex
Finish clean up of introduction

Treatments of evolving graphs vary in their generality and focus.
Kivel{\"a} \textit{et al.}~\cite{kabg14} treats time dependence as a special
case of families of graphs with multiple interrelationships.
Others like Flajolet \textit{et al.}~\cite{fkp89} use time to index the family
of related graph, but are not concerned with explicit time-dependent processes.

Our model can also be seen as a special case of
the multi-layer networks discussed by Kivel{\"a} \textit{et al.}~\cite{kabg14}, where
there is only one set of layers for time.

%XXX is the first sentence of this restatement true?
Previous work on evolving graphs tends to focus on special cases, such as
random graphs~\cite{fkp89} or sequences of graphs whose successive differences
are small~\cite{bkmu12}.  use the term to mean the evolution of random graphs.
Bahmani et al. \cite{bkmu12} use the term to mean a sequence of graphs where
successive differences are small.

Here, we are interested in extensions of graph algorithms that generalize for evolving graphs that work in full
generality.

%ARE THIS RELATED TO THE PREVIOUS WORK HOW
Work in this area include Despite the fact that graph metrics \cite{ntmm13,tmml09,tsmm09} and
various centrality measures like Katz centrality~\cite{gphe11,grihig13},
betweenness \cite{alhi15}, and closeness \cite{tmml10})
have been generalized for evolving graphs,
a BFS for evolving graphs, which we shall call it \emph{evolving-graph BFS},
was never explicitly described or verified.
%XXX THIS LEADS US TO CONSIDER ALGEBRAIC FORMULATIONS ALSO SINCE...

In this paper, we present a generalization of the breadth first search (BFS)
algorithm to evolving graphs. BFS is arguably the most fundamental graph search
algorithm, which leads to other graphical computations
such as the minimum spanning tree and shortest path \cite{ckrs09}.
We also demonstrate how BFS on evolving graphs can be expressed in the language
of linear algebra, and how our formulation helps analyze citation networks.

>>>>

Treatments of evolving graphs vary in their generality and focus.
Kivel\"a \textit{et al.}~\cite{kabg14} treats time dependence as a special
case of families of graphs with multiple interrelationships.
Others like Flajolet \textit{et al.}~\cite{fkp89} use time to index the family
of related graphs, but are not concerned with explicit time-dependent processes.
Yet others focus on incremental updates to large graphs~\cite{bkmu12}.
Here, we describe evolving graphs as a time-ordered sequence of graphs, similar
to the study of metrized graphs by Tang and coworkers~\cite{ntmm13,tmml09,tsmm09,tmml10}
the dynamics of communities by Grindrod, Higham and coworkers~\cite{gphe11,grihig13}.

The game described at the beginning can be encoded in terms of analyzing the
microscopic dynamics of an evolving graph. The spread of information to the
winner can be described in terms of discrete paths through the evolving graph,
and can be interpreted naturally as a graph traveral problem. Traversals of an
ordinary (static) graph may be computed using well known methods such as the
breadth-first search (BFS). However, a proper description of the dynamics
requires paths that also move monotonically forward in time, thus necessating
a generalization of the ordinary BFS algorithm. An informal description of such
a generalization was provided by \cite{tmml09}.

In this paper, we present a complete description of the BFS algorithm for evolving graphs.
The algebraic formulation of this algorithm makes explicit several subtleties in the correct
determination of relevant trajectories, which we term \textit{temporal paths},
which in turn can have important consequences for the computation and interpretation
of quantities which can be expressed as sums over paths, e.g.\ Katz centrality.
